### 2021-02-09 final_report/include/frontmatter/Titlepage.tex

% CREATED BY DAVID FRISK, 2016
% MODIFIED BY JAKOB JARMAR, 2016
% A few changes by Birgit Grohe, 2017 and \the\year
% Adjustments with the help of Gustav Örtenberg 2019

% COVER PAGE
\begin{titlepage}
\newgeometry{top=3cm, bottom=3cm,
			left=2.25 cm, right=2.25cm}	% Temporarily change margins

% Cover page background 
\AddToShipoutPicture*{\backgroundpic{-4}{56.7}{figure/auxiliary/frontpage_gu_eng_vec_m2.pdf}}
\addtolength{\voffset}{2cm}

% Cover picture (replace with your own or delete)		
\begin{figure}[H]
\centering
\vspace{1cm}	% Adjust vertical spacing here
\includegraphics[width=1\linewidth]{figure/cover/pyrgi.png}
\end{figure}

% Cover text
\mbox{}
\vfill
\renewcommand{\familydefault}{\sfdefault} \normalfont % Set cover page font

\textbf{\Huge \multiLineTitle{0.2cm}} 
\\[0.5cm]

%{\Large A Subtitle that can be Very Much Longer if Necessary}\\[0.5cm]

Master's thesis in Computer Science\setlength{\parskip}{1cm}

{\Large Arianna Masciolini} \setlength{\parskip}{2.9cm}

Department of Computer Science and Engineering \\
\textsc{Chalmers University of Technology} \\
\textsc{University of Gothenburg} \\
Gothenburg, Sweden \the\year

\renewcommand{\familydefault}{\rmdefault} \normalfont % Reset standard font
\end{titlepage}

% BACK OF COVER PAGE (BLANK PAGE)
\newpage
\restoregeometry
\thispagestyle{empty}
\mbox{}

% TITLE PAGE
\newpage
\thispagestyle{empty}
\begin{center}
	\textsc{\large Master's thesis \the\year}\\[4cm]		% Report number is currently not in use
	\textbf{\Large \multiLineTitle{0.2cm}} \\[1cm]
	{\large Arianna Masciolini}

	\vfill	
	% Logotype on titlepage	
	\begin{figure}[H]
	\centering
	% Remove the following line to remove the titlepage logotype
	\includegraphics[width=0.25\pdfpagewidth]{figure/auxiliary/ChGULogoHog.pdf}
	\end{figure}	\vspace{5mm}

	Department of Computer Science and Engineering\\
	%\emph{Division of Division name}\\
	%Name of research group (if applicable)\\
	\textsc{Chalmers University of Technology} \\
	\textsc{University of Gothenburg} \\
	Gothenburg, Sweden \the\year \\
\end{center}

% IMPRINT PAGE (BACK OF TITLE PAGE)
\newpage
\thispagestyle{plain}
\vspace*{4.5cm}
\oneLineTitle\\
Arianna Masciolini \setlength{\parskip}{1cm}

%\copyright ~ NAME FAMILYNAME, \the\year. \setlength{\parskip}{1cm}

Supervisor: Aarne Ranta, Department of Computer Science and Engineering\\
Examiner: Carl-Johan Seger, Department of Computer Science and Engineering\setlength{\parskip}{1cm}

Master's Thesis \the\year\\	% Report number currently not in use 
Department of Computer Science and Engineering\\
%Division of Division name\\
%Name of research group (if applicable)\\
Chalmers University of Technology and University of Gothenburg\\
SE-412 96 Gothenburg\\
Telephone +46 31 772 1000 \setlength{\parskip}{0.5cm}

\vfill
Cover: Reproduction of the Pyrgi Tablets, a bilingual Etruscan-Phoenician dedicatory text. Courtesy of the Museo Nazionale Etrusco di Villa Giulia.

Typeset in \LaTeX \\
%Printed by [Name of printing company]\\
Gothenburg, Sweden \the\year