### 2021-02-09 final_report/include/backmatter/Appendix_3.tex
added part about depconfigs

\chapter{Dependency configurations} \label{c}
This appendix provides the complete list of the abstract dependency configurations used for the UD-to-GF conversion of the alignments extracted by the CA module.

\section{Category annotation}
Category annotations describe a mapping between GF categories and UD \texttt{UPOS} tags (cf. Appendix \ref{a_pos}). \smallskip

\begin{verbatim}
    #cat A        ADJ primary
    #cat Adv      ADV primary
    #cat Det      DET primary
    #cat N        NOUN primary
    #cat PN       PROPN primary
    #cat Prep     ADP primary
    #cat Pron     PRON primary
    #cat V        VERB primary
    #cat Conj     CCONJ primary
\end{verbatim}

\section{Function annotations}
Function annotations, on the other hand, put GF functions in relation with UD \texttt{DEPREL}s (cf. Appendix \ref{a_lab}). 
The following list also serves the purpose of enumerating the types of concepts GF lexica automatically generated as described in Chapter \ref{ch5} consist of. \smallskip

\begin{verbatim}
    #fun AdjCN       amod head
    #fun AdvVP       head advmod
    #fun ComplV      head obj
    #fun DetCN       det head
    #fun AdvCN       head advmod
    #fun PredVP      nsubj head
    #fun PrepNP      case head
    #fun PrepCN      head nmod
    #fun FlatPN      head flat -- PN -> PN -> PN
    #fun CompoundN   compound head
    #fun PrepPP      head obl
\end{verbatim}